Binning Strategies and Related Loss for Large Data

### 2015-03-11 Paper.Rnw
aspect ratio

\begin{figure}[ht]
\centering
<<context, echo=false, fig.width=4, fig.height=4, out.width='.5\\textwidth'>>=
col <- c(rnorm(200, mean=0, sd=10), rnorm(10, mean=50), rnorm(200, mean=100, sd=10), rnorm(10, mean=50))
x <- c(rnorm(210, mean=0, sd=0.075), rnorm(210, mean=1, sd=0.075))
y <- c(rnorm(210, mean=0, sd=0.075), rnorm(210, mean=1, sd=0.075))

require(ggplot2)
qplot(x, y, colour=col, size=I(5)) + theme_bw()
@
\caption{\label{fig:context}Assess the colors in this scatterplot. Are the darkest spots in the cloud of lighter points as dark as the lightest spots in the cloud of darker points? Darker? Lighter? }
\end{figure}

@@

Frequency data consists of counts, which most commonly exhibit skew densities, i.e. there are usually a lot of cells with small cell counts and a few cells with extremely large counts. A log transformation therefore produces a more symmetric distribution of frequency information, increasing perceptual resolution. This is consistent with the Weber-Fechner law which suggests that increased stimulus intensity is perceptually mapped on the log scale \citep{sp}. Using a logarithmic mapping of frequency to the color or size aesthetic provides a more natural perceptual experience and simultaneously increases the perceptual resolution of the graph. The \textit{log frequency loss}, $L^{\log F}$, is defined as
%
\begin{eqnarray}\label{logfrequency}
L^{\log F} = \sum_{\ell=1}^{L} L_\ell^{\log F} = \sum_{\ell=1}^{L} (\log(c_\ell+1) - \log(b_c(c_\ell+1)))^2
\end{eqnarray}
%
Note that we have added one count to every cell before taking the logarithm to avoid problems with empty cells.

>>>>

\begin{figure}[ht]
\centering
<<context, echo=F, fig.width=6, fig.height=4.5, out.width='.5\\textwidth'>>=
col <- c(rnorm(200, mean=0, sd=10), rnorm(10, mean=50), rnorm(200, mean=100, sd=10), rnorm(10, mean=50))
x <- c(rnorm(210, mean=0, sd=0.075), rnorm(210, mean=1, sd=0.075))
y <- c(rnorm(210, mean=0, sd=0.075), rnorm(210, mean=1, sd=0.075))

require(ggplot2)
qplot(x, y, colour=col, size=I(5)) + theme_bw() + theme(aspect.ratio=1)
@
\caption{\label{fig:context}Assess the colors in this scatterplot. Are the darkest spots in the cloud of lighter points as dark as the lightest spots in the cloud of darker points? Darker? Lighter? }
\end{figure}

@@

Frequency data consists of counts, which most commonly exhibit skew densities, i.e.\ there are usually a lot of cells with small cell counts and a few cells with extremely large counts. A log transformation therefore produces a more symmetric distribution of frequency information, increasing perceptual resolution. This is consistent with the Weber-Fechner law which suggests that increased stimulus intensity is perceptually mapped on the log scale \citep{sp}. Using a logarithmic mapping of frequency to the color or size aesthetic provides a more natural perceptual experience and simultaneously increases the perceptual resolution of the graph. The \textit{log frequency loss}, $L^{\log F}$, is defined as
%
\begin{eqnarray}\label{logfrequency}
L^{\log F} = \sum_{\ell=1}^{L} L_\ell^{\log F} = \sum_{\ell=1}^{L} (\log(c_\ell+1) - \log(b_c(c_\ell+1)))^2
\end{eqnarray}
%
Note that we have added one count to every cell before taking the logarithm to avoid problems with empty cells.
